When two-point correlation function fails to describe a structure...

### 2025-01-10 paper.tex
Small style fix
Some another document class
Reformulate proof
Do not ommit mention of Dedekind zeta function

\documentclass[12pt, a4paper]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subfigure}

@@

\begin{document}
\title{When two-point correlation function fails to describe a structure...}
\author{Biba and Boba}
\maketitle

@@

@@

  For any $m \in \mathbb{Z} \setminus \{0, 1\}$ we have $m^2 > m$. Taking into
  account \cref{eq:coeff-equal} and letting any coefficient of $\tilde{f}$ be
  in set $\mathbb{Z} \setminus \{0, 1\}$ we get a contradiction with
  \cref{eq:squares-equal}. This contradiction proves this case.

@@

Now it must be clear how to find all the non-trivial collisions of $f$. To do so
one must follow \cref{alg:collisions}.
\begin{algorithm}
  \caption{An algorithm for finding non-trivial collisions of $f$ in the
    two-point correlation function.}
  \label{alg:collisions}
  \begin{algorithmic}[1]
    \Procedure{$Collisions$}{$f$}
    \State $fs \gets factor(f \in \mathbb{Z}[x])$
    \State $ps \gets filter(isPalindrome, fs)$
    \State $nps \gets fs \setminus ps$
    \If{$length(nps) \le 1$}
    \State \textbf{return} $f$ has only trivial collisions
    \EndIf
    \State $(np:nps') \gets nps$
    \State $f' \gets np \times \prod ps$
    \State $cs \gets \text{all possible combinations of $nps'$ or reciprocals}$
    \State \textbf{return} $map(f'\times, cs)$
    \EndProcedure
  \end{algorithmic}
\end{algorithm}
Here ``all possible combinations of $nps'$ or reciprocals'' needs
clarification. For each polynomial in $nps'$ either that polynomial or its
reciprocal is taken in a combination of factors $c$ and $cs$ is a list of all
possible $c$'s. If $f$ has $L$ non-palindrome factors as before, we get
$2^{L-1}$ polynomials (including $f$) which have the same two-point function as
$f$.

@@

Now we will prove a simple lemma which connects these results with collisions in
correlation functions.
\begin{theorem}
  \label{the:sequence}
  Let $A$ be a sufficiently long random sequence of 0's and 1's where each
  element is chosen randomly and independently with equal probability $1/2$ with
  exception of the first and the last elements which are 1's. Then, with
  probability at least
  \begin{equation*}
    P = 1 - 2\exp(-c(\log d)^\beta)
  \end{equation*}
  for some $c > 0$, $\beta > 1$ the sequence $A$ of the length $d+1$ does not
  have non-trivial collisions in the two-point correlation function.
\end{theorem}
\begin{proof}
  Let $f(x)$ be an image of $A$ under Z-transform and a change of variables
  $x = z^{-1}$. With probability at least $P$ it factors in $\mathbb{Z}[x]$ as
  $\Phi \tilde{f}$, where $\Phi$ is a product of cyclotomic polynomials and
  $\tilde{f}$ is irreducible. All cyclotomic polynomials with exception of
  $\Phi_1 = x - 1$ are palindromes. $\Phi_1$ does not divide $f$ because
  $f(1) = \sum\limits_{k=0}^{d-1} f_k \ge 2$. Therefore, $f$ has at most one
  non-palindrome factor and hence $A$ does not have non-trivial collisions.
\end{proof}

@@

  We have three possible cases of factorization of $f$:
  \begin{itemize}
  \item $f$ is irreducible.
  \item $f$ factors as $f(x, y) = g(x)h(y)$. In this case
    $\phi h = h(0) = const$ and \cref{lem:factors} cannot be applied. According
    to distribution of coefficients in $f$ the probability of this case is
    $P_{split} = 2^{D_1 + D_2 - D_1 D_2 - 1}$. For any fixed integer $D_2 > 1$
    we have $P_{split} = o(D_1^{-1})$. For $D_2 = 1$ we have $f = f_x$ and
    \cref{eq:prob-irr} is applicable directly to $f$.
  \item $f$ factors as $f(x, y) = g(x, y) h(x, y)$. According to conditions on
    the coefficients $\deg_x g > 0$ and $\deg_x h > 0$, therefore
    $\phi g \ne const$ and $\phi h \ne const$. We have
    $f_x = (\phi g) (\phi h)$. Applying \cref{lem:factors}, we can say that
    $f$ is irreducible with probability at least \cref{eq:main-result}.
  \end{itemize}
  Hence $f$ is irreducible with probability at least \cref{eq:main-result}. If
  $f$ is irreducible, there is no two or more non-palindrome factors of $f$,
  hence $A$ does not have non-trivial collisions.
\end{proof}
Surely, $D_1$ can be replaced with $D_2$ in \cref{eq:main-result} if
$A[0, D_2 - 1] = 1$. This theorem also can be proved for multidimensional arrays
by recursively applying the same argument.

@@

\newpage

\bibliographystyle{elsarticle-num}
\bibliography{paper}
\end{document}

>>>>

\documentclass[preprint,amsmath,amssymb,aps,pre,showkeys,showpacs]{revtex4-1}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subfigure}

@@

\begin{document}
\title{When two-point correlation function fails to describe a structure...}
\author{Biba}
\author{Boba}

@@

\maketitle

@@

  For any $m \in \mathbb{Z} \setminus \{0, 1\}$ we have $m^2 > m$. Taking into
  account \cref{eq:coeff-equal} and letting any coefficient of $\tilde{f}$ be
  in set $\mathbb{Z} \setminus \{0, 1\}$ we get a contradiction with
  \cref{eq:squares-equal}. This contradiction proves the case $n = 2$.

@@

Now it must be clear how to find all the non-trivial collisions of $f$. To do so
one must follow \cref{alg:collisions}.
\begin{algorithm}[H]
  \caption{An algorithm for finding non-trivial collisions of $f$ in the
    two-point correlation function.}
  \label{alg:collisions}
  \begin{algorithmic}[1]
    \Procedure{$Collisions$}{$f$}
    \State $fs \gets factor(f \in \mathbb{Z}[x])$
    \State $ps \gets filter(isPalindrome, fs)$
    \State $nps \gets fs \setminus ps$
    \If{$length(nps) \le 1$}
    \State \textbf{return} $f$ has only trivial collisions
    \EndIf
    \State $(np:nps') \gets nps$
    \State $f' \gets np \times \prod ps$
    \State $cs \gets \text{all possible combinations of $nps'$ or reciprocals}$
    \State \textbf{return} $map(f'\times, cs)$
    \EndProcedure
  \end{algorithmic}
\end{algorithm}
Here ``all possible combinations of $nps'$ or reciprocals'' needs
clarification. For each polynomial in $nps'$ either that polynomial or its
reciprocal is taken in a combination of factors $c$ and $cs$ is a list of all
possible $c$'s. If $f$ has $L$ non-palindrome factors as before, we get
$2^{L-1}$ polynomials (including $f$) which have the same two-point function as
$f$.

@@

Now we will prove a simple lemma which connects these results with collisions in
correlation functions.
\begin{theorem}
  \label{the:sequence}
  Let $A$ be a sufficiently long random sequence of 0's and 1's where each
  element is chosen randomly and independently with equal probability $1/2$ with
  exception of the first and the last elements which are 1's. Then, with
  probability at least
  \begin{equation*}
    P = 1 - 2\exp(-c(\log d)^{\beta-2})
  \end{equation*}
  for some $c > 0$, $\beta > 3$ the sequence $A$ of the length $d+1$ does not
  have non-trivial collisions in the two-point correlation function if $\zeta_K$
  has no zeros $\rho$ with $|1 - \rho| < (\log d)^\alpha/d$, $\alpha > \beta$
  for all $K = \mathbb{Q}(a)$ where $a$ is a root of $f(x)$ which is an image of
  $A$ under Z-transform and a change of variables $x = z^{-1}$.
\end{theorem}
\begin{proof}
  With probability at least $P$, $f \in \mathbb{Z}[x]$ factors as
  $\Phi \tilde{f}$, where $\Phi$ is a product of cyclotomic polynomials and
  $\tilde{f}$ is irreducible. All cyclotomic polynomials with exception of
  $\Phi_1 = x - 1$ are palindromes. $\Phi_1$ does not divide $f$ because
  $f(1) = \sum\limits_{k=0}^{d-1} f_k \ge 2$. Therefore, $f$ has at most one
  non-palindrome factor and hence $A$ does not have non-trivial collisions.
\end{proof}

@@

  We have three possible cases of factorization of $f$:
  \begin{itemize}
  \item $f$ is irreducible.
  \item $f$ factors as $f(x, y) = g(x)h(y)$. In this case
    $\phi h = h(0) = const$ and \cref{lem:factors} cannot be applied. According
    to distribution of coefficients in $f$ the probability of this case is
    $P_{split} = 2^{D_1 + D_2 - D_1 D_2 - 1}$. For any fixed integer $D_2 > 1$
    we have $P_{split} = o(D_1^{-1})$. For $D_2 = 1$ we have $f = f_x$ and
    \cref{eq:prob-irr} is applicable directly to $f$.
  \item $f$ factors as $f(x, y) = g(x, y) h(x, y)$. According to conditions on
    the coefficients $\deg_x g > 0$ and $\deg_x h > 0$, therefore
    $\phi g \ne const$ and $\phi h \ne const$. We have
    $f_x = (\phi g) (\phi h)$. Applying \cref{lem:factors}, we can say that this
    is the case with probability less that $1 - P$.
  \end{itemize}
  Hence $f$ is irreducible with probability at least \cref{eq:main-result}. If
  $f$ is irreducible, there is no two or more non-palindrome factors of $f$,
  hence $A$ does not have non-trivial collisions.
\end{proof}
Surely, $D_1$ can be replaced with $D_2$ in \cref{eq:main-result} if
$A[0, D_2 - 1] = 1$. This theorem also can be proved for multidimensional arrays
by recursively applying the same argument.

@@

\bibliography{paper}
\end{document}
